Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-16 paper/sections/proof-reconstruction.tex
[ paper ] fixing missing types, dots.

Therefore, for a general recursive function $f: \Prop \to \Prop$
with termination problems, we states a new function $f^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must done by
using induction on the \Nat type structure.

@@

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
&unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\
&unshunt_{1}(φ, \suc(n)) =\\
&\hspace{3mm}\begin{cases}
unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
  &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
unshunt_{1}(φ₁ ⇒ φ₂, n)\\
∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{split}
\end{align}

@@

In the case of the $unshunt_0$ function in
Eq.~(\ref{eq:unshunt-definition}), we define the complexity function
$unshunt_{cm}: \Prop → \NN$ for our $unshunt_{1}$
presented in Eq.~(\ref{eq:unshunt-rank-definition}).

@@

Finally, following the technique mentioned, we get a curated version of
$unshunt_{0}$ function presented in the following definition.

@@

Each proof of a subgoal is a refutation, thereby each proof assumes
the negation of its subgoal. The \negate rule
assumes the negation of such subgoals. This rules always
appears after apply the \strip inference to the goal.

@@

The \conjunct rule extracts from a
conjunction one of its conjuncts; it is a generalization of the
projection rules for the conjunction connective as the
following \TSTP excerpt shows.

@@

\begin{definition}[conjunct]
\begin{align}
  \begin{split}
  &conjunct : \Prop \to \Prop \to \Prop\\
  &conjunct(φ, ψ) =\\
  &\hspace{3mm}
  \begin{cases}
      ψ, &\text{if }φ ≡ ψ\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }ψ ≡ conjunct(φ₁, ψ)\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₁\text{ and }ψ ≡ conjunct(φ₂, ψ)\\
      φ, &\text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &reorder : \Prop → \Prop \to Prop\\
        &reorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{if } ψ ≡ ψ₁ ∧ ψ₂\text{, }reorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }reorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{if } ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct(φ, ψ), &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[conjunct$_{∨}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &conjunct_{∨} : \Prop → \Prop \to Prop\\
    &conjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{if }φ ≡ ψ\\
      ψ, &\text{if }reorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{if }ψ ≡ ψ₁ ∧ ψ₂, conjunct_{∨}(φ, ψ₁) ≡ ψ₁\\
        &\text{and } reorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₂, conjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₂, conjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[reorder$_{∧∨}$]
 \begin{align*}
      \begin{split}
      &reorder : \Prop \to \Prop \to \Prop\\
      &reorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{if } φ≡ψ\\
          ψ, &\text{if } ψ ≡ ψ₁ ∧ ψ₂, reorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }reorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{if }ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct_{∨}(φ, ψ) &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

@@

\begin{definition}[resolve]
\label{def:resolve}
 \begin{equation*}
    resolve(φ₁, φ₂, l, ψ) =
      rsol(reorder_{∨}(φ₁, l ∨ ψ) ∧ reorder_{∨} (φ₂, ¬ l ∨ ψ)).
  \end{equation*}
\end{definition}

@@

\begin{definition}[ndisj$_{∨}$]
  \label{def:ndisj-or}
  \begin{align*}
    \begin{split}
    &ndisj_{∨} : \NProp \to \NProp\\
    &ndisj_{∨}(φ) =
      \begin{cases}
        ⊤, &\text{if }φ ≡ ¬ φ₁ ∨ φ₂,\ φ₁ ∈_{∨} φ₂\\
        ⊤, &\text{if}φ ≡ ¬ φ₁ ∨ φ₂,\ ndisj_{∨}(φ₂) ≡ ⊤\\
        ¬ φ₁ ∨ ndisj_{∨}(φ₂), &\text{if }φ ≡ ¬ φ₁ ∨ φ₂\\
        ⊤, &\text{if }φ ≡ φ₁ ∨ φ₂,\ ¬φ₁ ∈_{∨} φ₂ \\
        ⊤, &\text{if }φ ≡ φ₁ ∨ φ₂,\ ndisj_{∨}(φ₂) ≡ ⊤\\
        φ₁ ∨ ndisj_{∨}(φ₂), &\text{if }φ ≡ φ₁ ∨ φ₂\\
        φ, &\text{otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

@@

\begin{lemma}[lem-dist$_{∨}$]
  \label{lem:lem-dist-or}
  $Γ ⟝ φ$ then $Γ ⟝ dist_{∨}(φ)$, where,
  \begin{align*}
      \begin{split}
      dist_{∨}(φ) =
        \begin{cases}
          dist_{∨}(φ₁ ∨ φ₂) ∧ dist_{∨}(φ₂ ∨ φ₃),
            &\text{if }φ ≡ (φ₁ ∧ φ₂) ∨ φ₃\\
          dist_{∨}(φ₁ ∨ φ₃) ∧ dist_{∨}(φ₁ ∨ φ₃),
            &\text{if }φ ≡ φ₁ ∨ (φ₂ ∧ φ₃)\\
          φ, &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{lemma}

\begin{lemma}[lem-dist]
  \label{lem:lem-dist}
  $Γ ⟝ φ$ then $Γ ⟝ dist(φ)$, where,
  \begin{align*}
      \begin{split}
      dist(φ) =
        \begin{cases}
          dist(φ₁) ∧ dist(φ₂),
            &\text{if }φ ≡ φ₁ ∧ φ₂\\
          dist_{∨}(dist(φ₁) ∨ dist(φ₂)),
            &\text{if }φ ≡ φ₁ ∨ φ₂\\
          φ, &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{lemma}

@@

\begin{align}
    \begin{split}
      &\textit{simpl} : \Prop \to \Prop \to \Prop \to \Prop\\
      &\textit{simpl}(φ₁, φ₂ ψ) =
      \begin{cases}
        ψ, &\text{if }φᵢ ≡ ψ,\text{ for some }i = 1, 2\\
        ψ, &\text{if }φ₁ ≡ φ₂ ⇒ ψ\\
      \end{cases}
    \end{split}
\end{align}

>>>>

Therefore, for a general recursive function $f: \Prop \to \Prop$
with termination problems, we states a new function
$f^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must be
done by using induction on the \Nat type structure.

@@

\begin{align}
\label{eq:unshunt-structured}
\begin{split}
&unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\%\hspace{10cm}\\
&unshunt_{1}(φ, \suc(n)) =\\
&\hspace{3mm}\begin{cases}
unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
  &\text{if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
unshunt_{1}(φ₁ ⇒ φ₂, n)\\
∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
   &\text{if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{split}
\end{align}

@@

We define the $unshunt_{cm}: \Prop → \NN$ function as
the complexity measure for $unshunt_0$ function.

@@

Finally, following the technique mentioned, we get a curated version of
$unshunt_{0}$ function presented in Def.~\ref{def:unshunt}.

@@

Each proof of a subgoal is a refutation, thereby each proof assumes
the negation of its subgoal. The \negate rule
assumes the negation of such subgoals. This rule
appears after applying the \strip inference to the goal.

@@

The \conjunct rule extracts from a
conjunction one of its conjuncts. This rule is a generalization of the
projection rules for the conjunction connective as the
following \TSTP excerpt shows us.

@@

\begin{definition}[conjunct]
\begin{align}
  \begin{split}
  &conjunct : \Prop \to \Prop \to \Prop\\
  &conjunct(φ, ψ) =
    \begin{cases}
        ψ, &\text{if }φ ≡ ψ\\
        ψ, &\text{if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ conjunct(φ₁, ψ)\\
        ψ, &\text{if }φ ≡ φ₁ ∧ φ₂\text{ , }ψ ≡ conjunct(φ₂, ψ)\\
        φ, &\text{otherwise.}
      \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &reorder : \Prop → \Prop \to \Prop\\
        &reorder_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{if } ψ ≡ ψ₁ ∧ ψ₂\text{, }reorder_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }reorder_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{if } ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct(φ, ψ), &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[conjunct$_{∨}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &conjunct_{∨} : \Prop → \Prop \to \Prop\\
    &conjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{if }φ ≡ ψ\\
      ψ, &\text{if }reorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{if }ψ ≡ ψ₁ ∧ ψ₂,\ conjunct_{∨}(φ, ψ₁) ≡ ψ₁,\\
        &\text{and } reorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₂,\ conjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{if }φ ≡ φ₁ ∧ φ₂,\ conjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[reorder$_{∧∨}$]
 \begin{align*}
      \begin{split}
      &reorder : \Prop \to \Prop \to \Prop\\
      &reorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{if } φ≡ψ\\
          ψ, &\text{if } ψ ≡ ψ₁ ∧ ψ₂,\ reorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }reorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{if }ψ ≡ ψ₁ ∧ ψ₂\\
          conjunct_{∨}(φ, ψ) &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

@@

\begin{definition}[resolve]
\label{def:resolve}
 \begin{align*}
 \begin{split}
    &resolve : \Prop \to \Prop \to \Prop \to \Prop \to \Prop\\
    &resolve(φ₁, φ₂, l, ψ) =
      rsol(reorder_{∨}(φ₁, l ∨ ψ) ∧ reorder_{∨} (φ₂, ¬ l ∨ ψ)).
  \end{split}
  \end{align*}
\end{definition}

@@

\begin{definition}[ndisj$_{∨}$]
  \label{def:ndisj-or}
  \begin{align*}
    \begin{split}
    &ndisj_{∨} : \NProp \to \NProp\\
    &ndisj_{∨}(φ) =
      \begin{cases}
        ⊤, &\text{if }φ ≡ ¬ φ₁ ∨ φ₂,\ φ₁ ∈_{∨} φ₂\\
        ⊤, &\text{if }φ ≡ ¬ φ₁ ∨ φ₂,\ ndisj_{∨}(φ₂) ≡ ⊤\\
        ¬ φ₁ ∨ ndisj_{∨}(φ₂), &\text{if }φ ≡ ¬ φ₁ ∨ φ₂\\
        ⊤, &\text{if }φ ≡ φ₁ ∨ φ₂,\ ¬φ₁ ∈_{∨} φ₂ \\
        ⊤, &\text{if }φ ≡ φ₁ ∨ φ₂,\ ndisj_{∨}(φ₂) ≡ ⊤\\
        φ₁ ∨ ndisj_{∨}(φ₂), &\text{if }φ ≡ φ₁ ∨ φ₂\\
        φ, &\text{otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

@@

\begin{lemma}[lem-dist$_{∨}$]
  \label{lem:lem-dist-or}
  $Γ ⟝ φ$ then $Γ ⟝ dist_{∨}(φ)$, where,
  \begin{align*}
      \begin{split}
      &dist_{∨} : \Prop \to \Prop\\
      &dist_{∨}(φ) =
        \begin{cases}
          dist_{∨}(φ₁ ∨ φ₂) ∧ dist_{∨}(φ₂ ∨ φ₃),
            &\text{if }φ ≡ (φ₁ ∧ φ₂) ∨ φ₃\\
          dist_{∨}(φ₁ ∨ φ₃) ∧ dist_{∨}(φ₁ ∨ φ₃),
            &\text{if }φ ≡ φ₁ ∨ (φ₂ ∧ φ₃)\\
          φ, &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{lemma}

\begin{lemma}[lem-dist]
  \label{lem:lem-dist}
  $Γ ⟝ φ$ then $Γ ⟝ dist(φ)$, where,
  \begin{align*}
      \begin{split}
      &dist : \Prop \to \Prop\\
      &dist(φ) =
        \begin{cases}
          dist(φ₁) ∧ dist(φ₂),
            &\text{if }φ ≡ φ₁ ∧ φ₂\\
          dist_{∨}(dist(φ₁) ∨ dist(φ₂)),
            &\text{if }φ ≡ φ₁ ∨ φ₂\\
          φ, &\text{otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{lemma}

@@

%TODO check canonicalize can reduce a conjunction into one smaller.
\begin{definition}[simpl]
\label{def:simpl}
\begin{align}
    \begin{split}
      &\textit{simpl} : \Prop → \Prop → \Prop\\
      &\textit{simpl}(φ₁, φ₂) =
      \begin{cases}
        ⊥,  &\text{ if }φᵢ ≡ ⊥,\\
        γ   &\text{ if }φᵢ ≡ φⱼ ⇒ γ\\
        γ   &\text{ if }φᵢ ≡ γ ⇒ ¬ φⱼ\\
        γ   &\text{ if }φᵢ ≡ φⱼ ⇔ γ \\
        γ   &\text{ if }φᵢ ≡ ¬ φⱼ ∨ γ\\
        ⊥   &\text{ if }conjunct(φᵢ, ¬ φⱼ)\\
        ⊥,  &\text{ if }canonicalize(φᵢ, ¬ φⱼ) ≡ ¬ φⱼ\\
        φ₁, &\text{ otherwise.}\\
      \end{cases}
    \end{split}
\end{align}
For $i = 1, 2$, $i\neq j$. We assume here some commutative properties for brevity.
\end{definition}
